---
title: "Analyse new dataset"
author: "Fabien Jossaud, Florent Chuffart"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---
 
  
```{r echo=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=TRUE, results="verbatim")
start_time = Sys.time()
source("common.R")
```

```{r params_default}
source("params_default.R")
if (!exists("gse_m")) {gse_m = gse}
```

In a conference to promote his article in NRG, Horvath introduces his epigenetic subject with the définition of a clock. 
For him, a clock is an accurate marker for chronological age. 
To precise his definition, he gives a definition of accuracy, saying a marker is accurate when you can have a $R^2$ > 0.8 for each independant data set including age between 0 and 100 yo.
So he gives different exemple to confirm his definition, like :
$$Telomere~size \sim age$$
$$Protein~expression~level \sim age$$
$$Gene~expression~level \sim age$$
$$DNAm \sim age$$
After giving all this examples of different clocks usable in different reviews, he says that the only model corresponding to his definition of clock and accuracy is his model : $$DNAm \sim age$$
Useful when it is his model.
So we decide to use epigenetic clocks to explains the chronological age. 

```{r building_indexes}
df = mreadRDS(paste0("df_preproc_",gse,".rds"))
idx_samples = rownames(df)
markers_start = grep("cg",colnames(df))[1]
idx_clinicals = colnames(df)[1:(markers_start-1)]
idx_cpg = colnames(df)[markers_start:ncol(df)]
```



```{r parameters}
idx_train = idx_test = rownames(df)
``` 

```{r loading train/test sets}
#Create Train and Test samples
Xtrain = mget_full_cpg_matrix(gse, idx_train, idx_cpg)
Ytrain = mget_df_preproc(gse)[idx_train,y_key]
Xtest = mget_full_cpg_matrix(gse, idx_test, idx_cpg)
Ytest = mget_df_preproc(gse)[idx_test,y_key]
``` 


```{r loading models}
litterature_models = mreadRDS("litterature_models.rds")
models = list(litterature_models$hannum_model_mc, litterature_models$horvath_model_mc)
if (!is.null(gse_m)) {
  models = c(mreadRDS(paste0("models_", gse_m, ".rds")), models)  
}
```

# Model Evaluation

```{r model evaluation, fig.width=12, fig.height=15, echo=FALSE, results="hide"}
for (covariate in covariates) {
  layout(matrix(1:20, 5), respect=TRUE)  
  for (m in models) {
    # layout(matrix(1:5, 1), respect=TRUE)
    dnamaging::plot_model_eval(m, df, covariate, Xtrain=Xtrain, Xtest=Xtest, Ytrain=Ytrain, Ytest=Ytest)
  }
}
```




# Session Information

```{r, results="verbatim"}
end_time = Sys.time()
print(paste0("Execution time for vignette : ", end_time - start_time))
sessionInfo()
```
