---
title: "Graphical comparison between iterative method and ewascombp method"
author: "Fabien Jossaud, Florent Chuffart"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---



```{r echo=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=FALSE, results="hide")
start_time = Sys.time()
source("common.R")
if (!exists("mopenxlsx_read.xlsx")) {mopenxlsx_read.xlsx = memoise::memoise(openxlsx::read.xlsx)}  
```



```{r params_default, echo=FALSE}
if (!exists("model_formula"))   { model_formula = "meth~age"            }
if (!exists("model_func_name")) { model_func_name = "modelcalllm"       }
if (!exists("newas"))           { newas = 1000000                           }
if (!exists("neighb"))          { neighb = 1000                         }
if (!exists("nb_run"))          { nb_run = 50                           }
if (!exists("pvals"))           { pvals = c(1e-05, 1e-10, 1e-20, 1e-30) }
if (!exists("gse_short"))           { gse_short="GSE42861" }

gse = paste0(gse_short,"_modelcalllm_meth~age_ewas", newas, "_nn", neighb)
study_filename = paste0("study_preproc_", gse,".rds") 
```


mkdir -p ./datashare/GSE42861/



rsync -auvP luke:~/projects/dnamaging/vignettes/models_r*_ewas*_nn1000_GSE42861_modelcalllm_meth~age_ewas*_nn1000.rds ~/projects/dnamaging/vignettes/.
rsync -auvP luke:~/projects/ewascombpr/vignettes/global_results_study_GSE42861_modelcalllm_meth~age_ewas*_nn1000.rds_modelcalllm_meth~age_1e-*.xlsx ~/projects/ewascombpr/vignettes/.
rsync -auvP luke:~/projects/ewascombpr/vignettes/study_GSE42861_modelcalllm_meth~age_ewas*_nn1000.rds ~/projects/dnamaging/vignettes/datashare/GSE42861/.
rsync -auvP luke:~/projects/dnamaging/vignettes/info_model_r*_ewas*_nn1000_GSE42861_modelcalllm_meth~age_ewas*_nn1000.rds ~/projects/dnamaging/vignettes/.


for (newas in c("200", "1000", "2000", "3000", "5000", "10000", "20000", "30000", "50000", "100000", "200000", "300000", "500000")) {
  print(paste0("****************newas: ", newas))
  source(knitr::purl("graphics_comparative_run.Rmd"))
}

for (newas in c("200", "1000", "2000", "3000", "5000", "10000", "20000", "30000", "50000", "100000", "200000", "300000", "500000")) {
  print(paste0("****************newas: ", newas))
  rmarkdown::render("graphics_comparative_run.Rmd", output_file=paste0("graphics_comparative_run_ewas", newas, ".html"))
}



```{r run comparative function, eval=TRUE}
recap_it_DMR = function(newas=1000000, neighb=1000, run=0, pval=1e-30, study_filename, gse, gse_short) { 
	foo = mreadRDS(paste0("./datashare/", gse_short, "/", study_filename))
	platform = foo$platform
	
	foo = read.table("~/projects/datashare/genomes/Homo_sapiens/UCSC/hg19/hg19.chrom.sizes")[-1,]
	xshift = c(0,cumsum(as.numeric(foo[,2]))[-nrow(foo)]) ; names(xshift) = foo[,1] 



	probes_it = list(probes = c(), it = c(), chr = c(), pos = c(), pos_tot = c())
  	rds_model_file = paste0("models_r",run,"_", gse,".rds")
  	if (!file.exists(rds_model_file)) {print(paste0("no file ", rds_model_file)) ; return()}
  	for (i in 0:run) { 
  		rds_model_file = paste0("models_r",i,"_", gse, ".rds")
  		print(rds_model_file)
		modele = mreadRDS(rds_model_file)
		modele = modele[[2]]$coeff$probes
		probes_it$probes = c(probes_it$probes,modele)
		probes_it$it = c(probes_it$it, rep(i, length(modele)))
		probes_it$chr = c(probes_it$chr,platform[modele,"chr"])
		probes_it$pos = c(probes_it$pos,platform[modele,"pos"])
	}
	probes_it$pos_tot = c(probes_it$pos_tot, probes_it$pos + xshift[probes_it$chr])	
	saveRDS(probes_it, paste0("meth_it_run",run,"_ewas", newas, "_nn", neighb, "_", gse, ".rds"))



  ewascombpr_global_results_file = paste0("~/projects/ewascombpr/vignettes/global_results_", study_filename, "_modelcalllm_meth~age_", pval, ".xlsx")
  print(ewascombpr_global_results_file)
	tab_DMR = mopenxlsx_read.xlsx(ewascombpr_global_results_file)
	tab_DMR = as.data.frame(tab_DMR)
	
	tab_comp = tab_DMR[,c("X.chrom","start","end","len","n_probes","probes")]
	
	# Probes de la méthode itérative présente dans chaque DMR
	probes_it_in_DMR = c()
	nb_probes_it_in_DMR = c()
	for (i in 1:nrow(tab_DMR)) {
		tmp_probes_it_in_DMR = probes_it$probes[probes_it$pos_tot <= tab_comp[i,"end"] + xshift[tab_comp[i,"X.chrom"]] & probes_it$pos_tot >= tab_comp[i,"start"] + xshift[tab_comp[i,"X.chrom"]]]
		tmp_nb_probes_it_in_DMR = length(tmp_probes_it_in_DMR)
		tmp_probes_it_in_DMR = paste(tmp_probes_it_in_DMR, collapse = ";")
		probes_it_in_DMR = c(probes_it_in_DMR, tmp_probes_it_in_DMR)
		nb_probes_it_in_DMR = c(nb_probes_it_in_DMR, tmp_nb_probes_it_in_DMR)
	}
	tab_comp = cbind.data.frame(tab_comp,probes_it_in_DMR,nb_probes_it_in_DMR)
	
	# Probes des DMR qui ne sont pas dans la méthode itérative 
	probes_DMR_not_in_it = c()
	for (i in 1:nrow(tab_comp)) {
		DMR_probes = strsplit(tab_comp[i,"probes"], split = ";")
		it_probes = strsplit(tab_comp[i,"probes_it_in_DMR"], split = ";")
		tmp_DMR_not_in_it = DMR_probes[!(DMR_probes %in% it_probes)]
		tmp_DMR_not_in_it = paste(unlist(tmp_DMR_not_in_it), collapse = ";")
		probes_DMR_not_in_it = c(probes_DMR_not_in_it,tmp_DMR_not_in_it) 
	}
	tab_comp = cbind.data.frame(tab_comp,probes_DMR_not_in_it)
	
	# pourcentage de probes de la méthode it présente et non presente dans chaque DMR	
	pct_probes_DMR_in_it = tab_comp[,"nb_probes_it_in_DMR"] / tab_comp[,"n_probes"] *100
	pct_probes_DMR_not_in_it = 100 - pct_probes_DMR_in_it
	tab_comp = cbind.data.frame(tab_comp,pct_probes_DMR_in_it, pct_probes_DMR_not_in_it)	
  saveRDS(tab_comp, paste0("tab_comp_run",run,"_ewas", newas, "_nn", neighb, "_pval",pval,"_", gse, ".rds"))
  # WriteXLS::WriteXLS(tab_comp, paste0("tab_comp_run",run,"_ewas", newas, "_nn", neighb, "_pval",pval,"_", gse, ".xlsx"), verbose=FALSE, row.names=FALSE, AdjWidth=TRUE, BoldHeaderRow=TRUE, FreezeCol=1)
	
	# Recap global 	
  probes_DMR = unlist(strsplit(tab_comp[,"probes"], split = ";"))
	probes_it_in_DMR = unlist(strsplit(tab_comp[,"probes_it_in_DMR"],split=";"))
	probes_it_not_in_DMR = probes_it$probes[!(probes_it$probes %in% probes_it_in_DMR)]
	nb_probes_it_in_DMR = length(probes_it_in_DMR)
	nb_probes_it_not_in_DMR = length(probes_it_not_in_DMR)
	pct_nb_probes_it_not_in_DMR = nb_probes_it_not_in_DMR / length(probes_it$probes) *100
	pct_gl_it_in_DMR = length(probes_it_in_DMR) / length(probes_DMR) *100
	pct_mean_it_in_DMR = mean(tab_comp[,"pct_probes_DMR_in_it"])
	
	stats_pres_it = data.frame(nb_probes_it_not_in_DMR, nb_probes_it_in_DMR)
	saveRDS(stats_pres_it, paste0("stats_pres_it_run",run,"_ewas", newas, "_nn", neighb, "_pval", pval,"_", gse, ".rds"))
	
	# Annotations probes not in DMR 
	
#	annot_it_not_in_DMR = data.frame(probes = probes_it_not_in_DMR,chr = probes_it$chr[!(probes_it$probes %in% probes_it_in_DMR)],it = probes_it$it[!(probes_it$probes %in% probes_it_in_DMR)],pos = probes_it$pos[!(probes_it$probes %in% probes_it_in_DMR)],pos_tot = probes_it$pos_tot[!(probes_it$probes %in% probes_it_in_DMR)])
#	WriteXLS::WriteXLS(annot_it_not_in_DMR, paste0("annot_it_not_in_DMR_run",run,"_pval",pval,".xlsx"), verbose=TRUE, row.names=FALSE, AdjWidth=TRUE, BoldHeaderRow=TRUE, FreezeRow=1, FreezeCol=1)
	
	print(paste0("run ", run, " terminated"))
	
	return()
	
}

for (pval in pvals) {
	for (run in (0:nb_run)) {
    print(paste0(pval," ", run))
			if(!file.exists(paste0("stats_pres_it_run",run,"_ewas", newas, "_nn", neighb, "_pval", pval,"_", gse, ".rds"))){
				recap_it_DMR(newas, neighb, run, pval, study_filename, gse)
			}
	}
}

```
# RMSE


```{r RMSE/nb_probes for iterative method}

####### RMSE nb_pb compare ######

## recover data 

x = c(0:nb_run)
rmse_bs = c()
rmse_glm = c()
nb_pb_bs = c()
nb_pb_glm = c()
for (run in 0:nb_run) {
	#tmp_model_bs = readRDS(paste0("info_model_r",run,"_ewas", newas, "_nn", neighb, "_", gse, ".rds"))
	tmp_model_bs = readRDS(paste0("info_model_r",run,"_", gse, "_modelcalllm_meth~age_ewas", newas, "_nn", neighb, ".rds"))
	tmp_rmse_bs = tmp_model_bs$gender.bootstrap.RMSE
	tmp_nb_pb_bs = tmp_model_bs$gender.bootstrap.nb_probes_mod
	rmse_bs = c(rmse_bs,tmp_rmse_bs)
	nb_pb_bs = c(nb_pb_bs, tmp_nb_pb_bs)
	
#	tmp_model_glm = readRDS(paste0("info_model_r",run,"_glm_ewas", newas, "_nn", neighb, "_", gse, ".rds"))
#	tmp_rmse_glm = tmp_model_glm$gender.elasticnet.RMSE
#	tmp_nb_pb_glm = tmp_model_glm$gender.elasticnet.nb_probes_mod
#	rmse_glm = c(rmse_glm,tmp_rmse_glm)
#	nb_pb_glm = c(nb_pb_glm, tmp_nb_pb_glm)
}

layout(1, respect=TRUE)

## add extra space to right margin of plot within frame
par(mar=c(5, 4, 4, 6) + 0.1)
## Plot first set of data and draw its axis
plot(x, rmse_bs, xlab="iteration", ylab="RMSE", 
   type="b", ylim=c(0,12), lty=1, col=1, main=paste0("RMSE~it newas=",newas), las=1)
abline(v=10, h=6, lty=2, col="grey")

## Allow a second plot on the same graph
par(new=TRUE)
## Plot the second plot and put axis scale on right
plot(x, nb_pb_bs,  xlab="", ylab="", ylim=c(0,200), 
    axes=FALSE, type="b", col="red", lty=1)
## a little farther out (line=4) to make room for labels
mtext("#probes",side=4,col="red",line=4) 
axis(4, col="red",col.axis="red",las=1)
```

# nb_probes

``` {r plot comparative graph DMR IT}
# FOR BS 
for (pval in pvals) {
	global_stat = c()
	for (run in (0:nb_run)) {
		tmp_stat = readRDS(paste0("stats_pres_it_run",run,"_ewas", newas, "_nn", neighb, "_pval", pval,"_", gse, ".rds"))
		global_stat = rbind(global_stat,tmp_stat)
	}
	rownames(global_stat) = c(0:nb_run)


	tab_DMR = mopenxlsx_read.xlsx(paste0("~/projects/ewascombpr/vignettes/global_results_", study_filename, "_modelcalllm_meth~age_", pval, ".xlsx"))
	tab_DMR = as.data.frame(tab_DMR)
	probes_DMR = unlist(strsplit(tab_DMR[,"probes"], split = ";"))
	nb_tot_DMR = length(probes_DMR)

  layout(matrix(1:2,1), respect=TRUE)
  ylim=c(0,1500)  
	plot(x=rownames(global_stat), y=global_stat$nb_probes_it_not_in_DMR, type = "b", col= "red", lty=1, ylim=ylim, main=paste0("cumsum(#probes) newas=", newas), ylab="#probes", xlab="iteration")
	points(x=rownames(global_stat), y = global_stat$nb_probes_it_in_DMR, type = "b", col = "blue",lty=1)
	points(x=rownames(global_stat), y = global_stat$nb_probes_it_not_in_DMR + global_stat$nb_probes_it_in_DMR, type = "b", col = "black",lty=1)
  abline(v=10, h=1000, lty=2, col="grey")
	# FOR GLM

	#global_stat = c()
	#for (run in (0:nb_run)) {

	#	tmp_stat = readRDS(paste0("stats_pres_it_run",run,"_ewas", newas, "_nn", neighb, "_pval", pval,"_", gse, ".rds"))
	#	global_stat = rbind(global_stat,tmp_stat)
	
	#}
	#rownames(global_stat) = c(0:nb_run)

	#points(x = rownames(global_stat), y = global_stat$nb_probes_it_not_in_DMR, type = "b", col = "red",lty=2)
	#points(x = rownames(global_stat), y = global_stat$nb_probes_it_in_DMR, type = "b", col = "blue",lty=2)
	#points(x = rownames(global_stat), y = global_stat$nb_probes_it_not_in_DMR + global_stat$nb_probes_it_in_DMR, type = "b", col = "black",lty=2)

	abline(h=nb_tot_DMR,col = "cyan",lwd=3)
	abline(h=sum(tab_DMR[,"n_probes"]),col="orange",lwd=3)
  plot.new()
	legend("topleft", col = c("red","blue","black","cyan","orange"), lty=c(rep(1,3),1,1) , lwd=c(rep(1,3),3,3), legend = c("nb bs iterative probes not included in DMR pipeline", "nb bs iterative probes inluded in DMR pipeline", "total nb of bs iterative probes", "number of DMR probes", "number of significative DMR probes" ), cex=.7)

	# legend("topleft", col = c("red","blue","black","red","blue","black","cyan"), lty=c(rep(1,3),rep(2,3),1) , lwd=c(rep(1,6),3), legend = c("nb bs iterative probes not included in DMR pipeline", "nb bs iterative probes inluded in DMR pipeline", "total nb of bs iterative probes", "nb glm iterative probes not included in DMR pipeline", "nb glm iterative probes inluded in DMR pipeline", "total nb of glm iterative probes","number of DMR probes"))

}

```

```{r manhattan plot, eval = FALSE}

probes_it = readRDS(paste0("meth_it_run",nb_run,"_ewas", newas, "_nn", neighb, "_", gse, ".rds"))
# it_in_DMR = as.data.frame(openxlsx::read.xlsx(paste0("tab_comp_run",run,"_ewas", newas, "_nn", neighb, "_pval",pval,"_", gse, ".xlsx")))
it_in_DMR = mreadRDS(paste0("tab_comp_run",run,"_ewas", newas, "_nn", neighb, "_pval",pval,"_", gse, ".rds"))
it_in_DMR = unlist(strsplit(it_in_DMR$probes_it_in_DMR,split=";"))
it_in_DMR = it_in_DMR[!(is.na(it_in_DMR))]

plot(x= probes_it$pos_tot, y = probes_it$it, col = as.numeric(probes_it$probes %in% it_in_DMR) +1 )

```



```{r isolated probes find, eval=FALSE}
tab_probes = openxlsx::read.xlsx(paste0("~/projects/ewascombpr/vignettes/global_results_", study_filename,".xlsx"))
probes_unique = unlist(tab_probes$probes[tab_probes$nb_probes == 1])
names(probes_unique) = probes_unique

probes_bs = readRDS(paste0("meth_it_run",nb_run,"_ewas", newas, "_nn", neighb, "_", gse, ".rds"))

probes_unique_it = probes_bs$it[probes_bs$probes %in%  probes_unique]
names(probes_unique_it) = probes_bs$probes[probes_bs$probes %in% probes_unique]

#probes_unique_not_it_name = probes_unique[!(probes_unique %in%  probes_bs$probes)]
#probes_unique_not_it = rep(-1,length(probes_unique_not_it_name))
#names(probes_unique_not_it) = probes_unique_not_it_name
#probes_unique = c(probes_unique_it, probes_unique_not_it)
layout(matrix(1:2,1), respect=TRUE)

plot(0:100,cumsum(table(factor(probes_unique_it, levels = 0:100))),type="b",ylim= c(0,length(probes_unique)),main=paste0("Isolated probes by iteration for ewas ",newas))
abline(h=length(probes_unique),col ="blue",lwd=3)

plot.new()
legend("topleft", legend = c("Isolated probes by iteration", "total of isolated probes"), col = c("black","blue"), lty = 1, lwd = c(1,3), cex=.7)

```

```{r plot density, eval = FALSE}
# Density ppv all probes 

if(file.exists("ppv_all_probes.rds")){ ppv_all_probes = readRDS("ppv_all_probes.rds") } else {

foo = mreadRDS(paste0("./datashare/", gse_short, "/", study_filename))
platform = foo$platform
	
foo = read.table("~/projects/datashare/hg38.chrom.sizes")[1:24,]
xshift = c(0,cumsum(as.numeric(foo[,2]))[-nrow(foo)]) ; names(xshift) = foo[,1] 
	
all_probes_pos_tot = platform$pos + xshift[platform$chr] 


ppv_all_probes = c()
for (i in 1:length(all_probes_pos_tot)) {
	tmp_ppv = c()
	for (j in c(1:i-1,i+1:length(all_probes_pos_tot))) {
		tmp_ppv = c(tmp_ppv, abs(all_probes_pos_tot[i] - all_probes_pos_tot[j]))
	}
	ppv_all_probes = c(ppv_all_probes, min(tmp_ppv))
	print(i)
}

}

den_ap = density(ppv_all_probes)
bw = den_ap$bw


plot(density(ppv_all_probes, bw), col = "black", main = "Distance between interest probes comparison")

print("all probes done")

# Density k-ppv it probes glm


#probes_it = readRDS(paste0("meth_it_run",nb_run,"_ewas", newas, "_nn", neighb, "_", gse, ".rds"))

#ppv_glm = c()
#for (i in 1:length(probes_it$pos_tot)) {
#	tmp_ppv = abs(probes_it$pos_tot[i] - probes_it$pos_tot)[-i]
#	ppv_glm = c(ppv_glm, min(tmp_ppv))
#}

#lines(density(ppv_glm, bw), col = "red")

#print("glm done")

# Density k-ppv it probes bs

probes_it = readRDS(paste0("meth_it_run",nb_run,"_ewas", newas, "_nn", neighb, "_", gse, ".rds"))

ppv_bs = c()
for (i in 1:length(probes_it$pos_tot)) {
	tmp_ppv = abs(probes_it$pos_tot[i] - probes_it$pos_tot)[-i]
	ppv_bs = c(ppv_bs, min(tmp_ppv))
}

lines(density(ppv_bs, bw), col = "green")

print("bs done")

# Density ppv DMR

pval = 1e-30

tab_DMR = openxlsx::read.xlsx(paste0("~/projects/ewascombpr/vignettes/global_results_", study_filename, "_modelcalllm_meth~age_", pval, ".xlsx"))
tab_DMR = as.data.frame(tab_DMR)

foo = read.table("~/projects/datashare/hg38.chrom.sizes")[1:24,]
xshift = c(0,cumsum(as.numeric(foo[,2]))[-nrow(foo)]) ; names(xshift) = foo[,1] 
	
tab_DMR$tot_start = tab_DMR$start + xshift[tab_DMR$X.chrom] 
tab_DMR$tot_end = tab_DMR$end + xshift[tab_DMR$X.chrom]

ppv_DMR = c()

for (i in 1:nrow(tab_DMR)) {
	tmp_ppv = c()
	for (j in c(1:nrow(tab_DMR))[-i]) {
		if (tab_DMR$tot_start[i] < tab_DMR$tot_start[j]) {
			tmp_ppv = c(tmp_ppv, abs(tab_DMR$tot_end[i] - tab_DMR$tot_start[j]))
		} else {
			tmp_ppv = c(tmp_ppv, abs(tab_DMR$tot_start[i] - tab_DMR$tot_end[j]))
		}
	}
	ppv_DMR = c(ppv_DMR, min(tmp_ppv))
}

lines(density(ppv_DMR, bw),col = "blue")

print("DMR done")

legend("topleft", legend = c("distance of platform probes NN", "distance of DMRs NN", "distance of glm iterative probes NN", "distance of bs iterative probes NN"), col = c("black","blue","red","green"), lty = 1)

```

# exec_time
```{r}

####### RMSE nb_pb compare ######

## recover data 

x = c(0:nb_run)
exec_time = c()
for (run in 0:nb_run) {
	tmp_model_bs = readRDS(paste0("info_model_r",run,"_ewas", newas, "_nn", neighb, "_", gse, ".rds"))
	tmp_exec_time = as.numeric(tmp_model_bs$exec_time)
	exec_time = c(exec_time,tmp_exec_time)
	
#	tmp_model_glm = readRDS(paste0("info_model_r",run,"_glm_ewas", newas, "_nn", neighb, "_", gse, ".rds"))
#	tmp_rmse_glm = tmp_model_glm$gender.elasticnet.RMSE
#	tmp_nb_pb_glm = tmp_model_glm$gender.elasticnet.nb_probes_mod
#	rmse_glm = c(rmse_glm,tmp_rmse_glm)
#	nb_pb_glm = c(nb_pb_glm, tmp_nb_pb_glm)
}

layout(matrix(1:2,1), respect=TRUE)

## add extra space to right margin of plot within frame
par(mar=c(5, 4, 4, 6) + 0.1)

## Plot first set of data and draw its axis
plot(x, exec_time, xlab="iteration", type="b", ylim=c(0,15), lty = 1, col="black", main=paste0("Comparison GLM/BS method for ewas=",newas), ylab="exec_time (min)")
```
