---
title: "Code Hannum 2013"
---
title: "Pipeline Hannum"
author: "Fabien Jossaud, Florent Chuffart"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---


```{r include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=TRUE, results="verbatim")
```

```{r params}

if (!exists(gse)) {gse = "GSE41037"}
if (!exists(y_key)) {y_key = "age"}

```

```{r batch, eval=FALSE} 

output_file = paste0("05_Pipeline_", gse, ".Rmd")

if (!file.exists(output_file)) {
    rmarkdown::render("05_Pipeline.Rmd", output_file=output_file)    
}

``` 

# Call data (Need to be modified to generalize and make it at a variable)

```{r call data}

s_filename = paste0("df_", gse, ".rds")

if (!exists("mreadRDS")) { mreadRDS = memoise::memoise(readRDS) }

if (!exists("dfs")) { 
  dfs = list()
}

if (!gse %in% names(dfs)) {
  dfs[[gse]] = mreadRDS(s_filename)  
}

df = dfs[[gse]] 

idx_samples = rownames(df)

# Check where begin markers 

markers_start = grep("cg",colnames(df))[1]

idx_clinicals = colnames(df)[1:(markers_start-1)]

idx_cpg = colnames(df)[markers_start:ncol(df)][1:10000]

## Cut Train/Test for prediction comparison. 

Ntrain = floor(nrow(df)/2)
set.seed(1)

idx_train = sample(rownames(df),Ntrain)
idx_test = setdiff(rownames(df),idx_train)


Train = df[idx_train,]
dim(Train)

Test = df[idx_test,]
dim(Test)



```

# Pretreatment

## Missing values

```{r Missing Values}
if (sum(is.na(df[,idx_cpg]))>0) {
  # Looking how much got more than 5% missing values
  # apply_func = epimedtools::monitored_apply
  if (! exists("apply_func")) {apply_func = apply}
  prop_na_row =  apply_func(is.na(df[,idx_cpg]), 1, sum) / ncol(df)
  idx_samples = names(prop_na_row)[prop_na_row<.05]
  prop_na_col =  apply_func(is.na(df[,idx_cpg]), 2, sum) / nrow(df)
  idx_cpg     = names(prop_na_col)[prop_na_col<.05]
  df = df[idx_samples, c(idx_clinicals, idx_cpg)]

  ## KNN - We use the function impute.knn of the package impute to replace remaining missing values with KNN method. 
  dknn = df
  dknnf = impute::impute.knn(df, k=10, maxp=5000, rng.seed=1)
  df = dknnf$data
}
```


# Outliers 

## ACP 

We first make an ACP to find outliers.

```{r ACP}
padj = 1
idx_samples
it=1
while(sum(padj < 0.2) != 0) {
  print(paste0("*******", it))
  it = it+1
  mat = as.matrix(df[idx_samples, idx_cpg])
  pca = prcomp(mat, scale=FALSE)
  v = pca$sdev * pca$sdev
  p = v / sum(v) * 100

  layout(matrix(1:6,2), respect=TRUE)
  # layout(matrix(1:2,1), respect=TRUE)
  barplot(p)

  for (i in 1:5) {
    j = i+1
    plot(pca$x[,i], pca$x[,j], xlab=paste0("PC", i, "(", signif(p[i], 3), "%)"), ylab=paste0("PC", j, "(", signif(p[j], 3), "%)"), pch=16)  
    # scale_factor = min(abs(c(min(c(pca$x[,i], pca$x[,j])), max(c(pca$x[,i], pca$x[,j])))))
    # scale_factor = min(abs(c(max(min(pca$x[,i]), min(pca$x[,j])), min(max(pca$x[,i]), max(pca$x[,j])))))
    # arrows(0,0,pca$rotation[,i]*scale_factor, pca$rotation[,j]*scale_factor, col="grey")
    # text(pca$rotation[,i]*scale_factor, pca$rotation[,j]*scale_factor, rownames(pca$rotation))
  }

  head(pca$x[,1:2])

  mean = mean(pca$x[,1])
  sd = sd(pca$x[,1])
  pval = pnorm(abs(pca$x[,1]), mean, sd, lower.tail=FALSE)
  padj = p.adjust(pval, method="BH")
  sum(padj <= 0.2)
  range(padj)  
  idx_samples = names(padj)[padj > 0.2]
}
```

# Model 

## Find Variables for model 

```{r Prepare Model}

Xtrain = Train[,idx_cpg]
Ytrain = Train[,y_key]

Xtest = Test[,idx_cpg]
Ytest = Test[,y_key]

``` 

## Finding Best Model
 
### Test Classic regression

```{r Classic Regression}

## Prepare data

Xtrain = data.matrix(Xtrain)

Xtest = data.matrix(Xtest)

## Cross validation for best lambda and alpha

alphas = seq(0.5,1,0.5)

stat = lapply(alphas,function(alpha){ 
  
  modelcv = glmnet::cv.glmnet(Xtrain,Ytrain, alpha = alpha, type.measure = "mse")
  lambdamin = modelcv$lambda.min
  cvrecap = data.frame(lambda = modelcv$lambda, RMSE = sqrt(modelcv$cvm), nbprobes = modelcv$nzero)
  ret = list(
    cvrecap = cvrecap, 
    lambdamin = lambdamin,
    # model = modelcv,
    alpha = alpha
  )
  return(ret)
  
  })

layout(matrix(1:2,1),respect = TRUE)

plot(0,0,col = 0, xlim = c(0,3), ylim = c(4,10))

lapply(stat, function(it) { 
  
  cvrecap = it$cvrecap
  points(cvrecap$lambda,cvrecap$RMSE)
  abline(v = it$lambdamin)
  })

plot(0,0,col = 0, xlim = c(0,3), ylim = c(0,500))

lapply(stat, function(it) { 
  
  cvrecap = it$cvrecap
  points(cvrecap$lambda,cvrecap$nbprobes)
  abline(v = it$lambdamin)
  })

## Plotting alpha and lambda.min predictions 

layout(matrix(1:4,2),respect = TRUE)
par(mfrow = c(2,2))
lapply(stat, function(it) {
  
  # it = stat[[1]]
  model = glmnet::glmnet(Xtrain,Ytrain, alpha = it$alpha, lambda = it$lambdamin, standardize = TRUE) 
                         
  predTr = predict(model,Xtrain)
  RMSETr = signif(sqrt(mean((Ytrain - predTr)^2, na.rm = TRUE)),3)
  plot(predTr,Ytrain,main = paste0("Pred Train set alpha = ",it$alpha,", RMSE = ", RMSETr))

  predTe = predict(model,Xtest)
  RMSETe = signif(sqrt(mean((Ytest - predTe)^2, na.rm = TRUE)),3)
  plot(predTe,Ytest,main = paste0("Pred Test set alpha = ",it$alpha,", RMSE = ", RMSETe))  
  
})
  
```

### Horvath Package

 
```{r methylclock}

## Prepare data 

XtestMC = t(Xtest)

XtrainMC = t(Xtrain)

## Hannum clock

### Train 

modelHOTr = methylclock::DNAmAge(XtrainMC,age=Ytrain,cell.count = FALSE, clocks = "Hannum")

RMSEHOTr = sqrt(mean((modelHOTr$age - modelHOTr$Hannum)^2))

plotAMARHOTr = methylclock::plotDNAmAge(modelHOTr$Hannum, modelHOTr$age, paste0("AMAR fonction DNAmAge with Hannum Method on Train Set, RMSE = ",RMSEHOTr))

plotAMARHOTr

### Test

modelHO = methylclock::DNAmAge(XtestMC,age=Ytest,cell.count = FALSE, clocks = "Hannum")

RMSEHO = sqrt(mean((modelHO$age - modelHO$Hannum)^2))

plotAMARHO = methylclock::plotDNAmAge(modelHO$Hannum, modelHO$age, paste0("AMAR fonction DNAmAge with Hannum Method on Test Set, RMSE = ",RMSEHO))

plotAMARHO

## Horvath clock

### Train

modelHOrTr = methylclock::DNAmAge(XtrainMC,age=Ytrain,cell.count = FALSE, clocks = "Horvath")

RMSEHOrTr = sqrt(mean((modelHOrTr$age - modelHOrTr$Horvath)^2))

plotAMARHOrTr = methylclock::plotDNAmAge(modelHOrTr$Horvath, modelHOrTr$age, paste0("AMAR fonction DNAmAge with Horvath Method on Train Set, RMSE = ",RMSEHOrTr))

plotAMARHOrTr

### Test 

modelHOr = methylclock::DNAmAge(XtestMC,age=Ytest,cell.count = FALSE, clocks = "Horvath")

RMSEHOr = sqrt(mean((modelHOr$age - modelHOr$Horvath)^2))

plotAMARHOr = methylclock::plotDNAmAge(modelHOr$Horvath, modelHOr$age, paste0("AMAR fonction DNAmAge with Horvath Method on Test Set, RMSE = ",RMSEHOr))

plotAMARHOr

## RMSE Comparison

RMSE_Tr = c(RMSECRTr,RMSEHOTr,RMSEHOrTr)
RMSE_Te = c(RMSECR,RMSEHO,RMSEHOr)
RMSE_compar = as.data.frame(rbind(RMSE_Tr,RMSE_Te))
colnames(RMSE_compar) = c("Selfmade Regression","Hannum method DNAmAge function","Horvath method DNAmAge function")
rownames(RMSE_compar) = c("Train","Test")



```
# Session Information

```{r, results="verbatim"}
sessionInfo()
```
